Texture Mapping 3D Models of Indoor Environments with Noisy Camera Poses

### 2013-05-31 masters/body/masters.tex
so much

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover its pose within an environment while simultaneously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment. Mounting such devices on a platform carried by
an ambulatory human provides unique advantages over static or
vehicular-based systems on wheels in terms of agility and portability,
but can also result in larger localization error
\cite{liu2010indoor}. As a result, common methods for texture mapping
generally produce poor results. In this paper, we present an approach
to texture mapping 3D models of indoor environments in the presence of
uncertainty and noise in camera poses. In particular, we consider data
obtained from a human-operated backpack system, detailed in Section
\ref{sec:dataAcquisitionSystem}.

\begin{figure}
  \centering
  \includegraphics[width=6in]{flowchart.jpg}
  \caption{The proposed texture mapping procedure ***this will
    change\\}
  \label{fig:flowchart}
\end{figure}

The overall block diagram for the proposed texture mapping procedure
is shown in Figure \ref{fig:flowchart}, where the number attached
attached to each box indicates the section in which the concept in the
box is explained in this paper ***will change. First, the geometry of
an environment is split into regions. For each region, we begin by
selecting a set of images that spans the entire region's surface with
high resolution imagery. We then use recovered noisy camera poses to project
these selected images onto the surface. These projections are rotated
and translated in 2D, in order to maximally align them with the
surface's geometry, as well as to each other, allowing us to handle
both errors in geometry as well as camera poses. For surfaces and
camera poses at arbitrary locations or orientations, generally
ceilings and floors, we propose a tile-based approach for sampling
high-resolution portions of images and compositing them into a
texture. In cases where cameras have consistently perpendicular
viewing angles to the surfaces under consideration, generally walls,
we demonstrate a superior approach based on a shortest-path
computation that leads to more seamless textures.

The remainder of the paper is organized as follows. Section
\ref{sec:dataAcquisitionSystem} describes our data acquisition system
as well as our region segmentation procedure. Section
\ref{sec:existingApproaches} covers existing approaches to image
stitching, and their performance on our datasets. Section
\ref{sec:simpleTextureMapping} explains how to downsample the set of
available images by selecting those with the best orientation and
distance from surfaces under consideration. Section
\ref{sec:2dAlignment} contains our approach towards 2D image
alignment, followed by Section \ref{sec:imageCompositing}, which
describes two methods of selecting and compositing images to create
the final texture. Sections \ref{sec:results} and \ref{sec:conclusion}
contain results and conclusions.

@@

A CAD model of the backpack-mounted system is shown in Figure
\ref{fig:backpack}. This backpack system has a number of laser range
scanners as well as 2 cameras facing left and right, each equipped
with fisheye lenses reaching an approximately 180$^{\circ}$ field of
view each and taking photos at a rate of 5 Hz \cite{liu2010indoor}. In
order to collect data, an operator wears the
backpack system and traverses the environment to be scanned, walking
in straight lines at a standard pace and making an effort to walk
parallel to all walls in the area. Multiple localization and
loop-closure algorithms are then applied to the raw data collected by
the onboard sensors \cite{chen2010indoor, kua2012loopclosure,
  liu2010indoor}, and the backpack is localized\footnote{In this
  paper, we use the terms localization and pose recovery
  interchangeably, in that they both refer to recovering position and
  orientation.}  over its data collection period. This involves
recovering the 6 degrees of freedom for the backpack itself as well as
the cameras rigidly mounted on it. Once this is complete, the data
from the laser range scanners is used to generate a 3D point cloud of
the surrounding environment. This point cloud is used to construct
either a low-resolution planar model, such as for usage in basic
navigation or energy analysis \cite{sanchez2012point,
  turnerfloorplane}, or a high-resolution, detailed triangular mesh,
for applications such as accurate remodeling or virtual walkthroughs
\cite{turnerwatertight}.  Regardless of resolution, size, or detail,
the model, consisting of geometric surfaces in 3D space, along with
the set of images captured by the backpack's cameras and their noisy
3D poses, are the input to our texture mapping problem.

\subsection{Geometry Partitioning}
\label{sec:geometryPartioning}
In order to effectively texture a 3D model, we first divide the model
into a set of regions, each to be textured independently. The purpose
of this is to ensure texture continuity along large uniform areas,
while allowing texture boundaries to fall along natural environmental
boundaries. Because textures may not match at such boundaries where
two regions meet, it is important to minimize the visibility of
boundaries. This is done by encouraging region boundaries to occur at
large sharp corners, where texture continuity is less important.

When working with low-resolution models, where surfaces generally
correspond to large planar features such as walls, ceilings, or
floors, and often meet at right angles, each planar surface in the
model is directly treated as a distinct region and textured
separately.

When working with a high-resolution model however, geometry often
represents large features as well as small details found in furniture,
plants, and other miscellaneous objects. For instance, in Figure
\ref{fig:deskGeom}(a), the geometry modeling the desk and binders
consists of many small triangles at various orientations. Texturing
each of these triangles independently makes little sense, as the
textures applied to each should match up well with the textures of its
neighbors in order to avoid any breaks in the overall image of the
desk and binders. On the other hand, the textures for the triangles
belonging to the wall behind need not match up with the textures on
the desk and binders, and so each group of triangles should be
designated to separate regions. This region grouping is performed by
designating all contiguous coplanar groups of triangles as different
regions. Any region that is less than $1 m^2$ in size is joined to its
largest neighboring region, such that the angle between them is under
$90^{\circ}$. An example of such a grouping is shown in
\ref{fig:deskGeom}(b).

Though most regions are not completely flat, we fit a plane to each
region for the purposes of texturing. This plane corresponds to the
largest contiguous section of coplanar triangles, as most regions are
characterized by small outcroppings from a largely planar surface,
such as objects on a table or features protruding from a wall. The 3d
vertices within a region are then projected onto its calculated plane,
and the 2D contour for these projected vertices is calculated.

Of course, the 2D contour of a region is not as accurate as its
original 3D geometry. Thus, we also create a k-d tree consisting of
the triangle primitives from the original mesh. Though textures are
generated upon the 2D planar surface, the original 3D geometry in the
k-d tree is referenced to ensure accuracy at various stages when
texturing.

At this point, we either have a series of planar regions, or planar
approximations for regions and an associated k-d tree. The task now is
to generate a texture for each region, though first some related work
will be discussed.

\section{Related Work}
\label{sec:relatedWork}
There are many existing approaches to stitching together multiple
images to produce a larger, seamless image \cite{szeliski2006image,
  agarwalapanoramas, wangmultipleviews, coorg1997matching,
  debevechybrid, bernardinimultiplescans}. Generally, parts of images
are matched to each other, either through direct comparisons at pixel
level, or through feature detection and matching. Images are then
transformed to maximize matches, often by computing homographies
between pairs of images, or by iteratively adjusting camera poses in 1
to 6 degrees of freedom.

@@

In prior work, we have experimented with integrating image stitching
with the iterative global localization algorithm used to recover the 6
degrees of freedom of the backpack acquisition system detailed in
Section \ref{sec:dataAcquisitionSystem} \cite{liu2010indoor}. When run
on long chains of images however, especially where features are
sparse, this approach produces distorted textures, as seen in Figure
\ref{fig:mosaic3D}(a). Furthermore, this approach is not closed-form,
and its iterative camera adjustment process over large datasets leads
to prohibitively long computation time.

The AutoStitch software package, based on research in the area of
panorama genoration, performs homography-based alignment as well, with
additional provisions that attempt to reduce drift and increase
overall efficiency \cite{panorama2d, autostitch}. Though AutoStitch
performs well in small areas, it often stiff induces drift in images,
such that global alignment to the environment geometry is
lost. Furthermore, its performance relies on the presence of a dense
set of features, and it has trouble aligning wall sections with even
short segments of bare texture. The example in Figure
\ref{fig:mosaic3D}(b) was generated after many rounds of manual
tuning; we have empirically found that areas with fewer visual
features or repeating texture patterns simply failed outright with
AutoStitch.

@@

The geometry of the texture mapping process for a region, as described
in Section \ref{sec:geometryPartitioning}, is shown in Figure
\ref{fig:projection}. Given a set of $M$ images to texture a target
plane, camera matrix $P_i$ for the $ith$ image transforms a 3D point
in the world coordinate system to a 2D point or pixel in image $i$'s
coordinates. A camera matrix $P_i$ is composed of the camera's
intrinsic parameters, containing focal length and image center, as
well as extrinsic parameters which specify the rotation and
translation of the camera's position in 3D world coordinates at the
time that image $i$ is taken. These extrinsic parameters are
determined by the backpack hardware and the corresponding localization
algorithms \cite{chen2010indoor, liu2010indoor, kua2012loopclosure}
and are noisy.

Since the backpack system takes pictures at a rate of 5 Hz, thousands
of images are available for texturing each surface in the 3D
model. Our objective in designing a texture mapping process is to
determine which of these images should be used, and where their
contents should map onto the final texture, in order to minimize any
visual discontinuities or seams that would suggest that the plane's
texture is not composed of a single continuous image. In the remainder
of this section, we propose an image subsampling procedure and use it
to derive a simple tile-based texture mapping procedure.

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.8in]{Projection.pdf}
    \caption{Surfaces to be textured are specified in 3D space by
      corners $C_i$. Images are related to each surface through the
      camera matrices $P_{1..m}$. }
    \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.8in]{scoringFunction.jpg}
    \caption{We minimize camera angle $\alpha$ and distance $d$ by
      maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
      \cdot \vec{n}$}
    \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
a straightforward approach for texturing is to discretize the target
plane into small square tiles, and choose an image to texture each
tile directly. We choose to work with rectangular units to ensure that
borders between any two distinct images in the final texture are
either horizontal or vertical. Since most environmental features
inside buildings are horizontal or vertical, any visible seams in the
texture intersect them minimally and are therefore less noticeable.

In order to select an image for texturing a tile $t$, we first gather
a list of candidate images that contain all four of its corners; we
can rapidly check this by projecting $t$ into each image using the
$P_i$ camera matrices. Furthermore, each candidate image must have
been taken at a time when its camera had a clear line-of-sight to $t$,
which can be determined using standard ray-polygon intersection tests
between the camera location, $t$, and every other surface
\cite{rayintersection}.

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since resolution decreases and camera pose errors become more
pronounced with distance, we wish to minimize the distance between
cameras and the surfaces they texture. Additionally, we desire images
that are projected perpendicularly, rather than obliquely, onto the
plane, maximizing the resolution and amount of useful texture
available in their projections, as well as minimizing any parallax
effects due to real-world geometry not accurately represented by the
digital 3D model. In other words, we wish to minimize the angle
between the tile's normal vector and the camera axis for images
selected for texturing that tile. These two criteria can be met by
maximizing the function $\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$
as shown in Figure \ref{fig:scoringFunction}, where $d$ is the
distance between the centers of a camera and a tile, and $\vec{n}$ and
$\vec{c}$ are the directions of the plane's normal and the camera axis
respectively. As Figure \ref{fig:compareAll}(a) demonstrates, this
approach, when used directly for texturing, results in image
boundaries with abrupt discontinuities between tiles, due to
misalignment between images, though most of it appears to be
reconcilable using 2D transforms.

While this procedure can be used to generate a texture, it is also
useful for obtaining a collection of images that can be used as a
starting point for our proposed texturing procedure, as explained in
Section \ref{sec:2dAlignment}. For our datasets, this process
generally selects around 10\% of all the images that could be used for
texturing the surface. This roughly 10:1 subsampling not only reduces
computational complexity of the remaining steps, but also selects the
most promising images for the remaining steps.

@@

\section{2D Image Alignment}
\label{sec:2dAlignment}
In this section, we describe our method for efficient and robust image
alignment. Rather than register all of our images in 3D, as many
state-of-the-art techniques for image stitching do, we instead select
a subset of images to be aligned in 2D. This subset corresponds to the
images selected by the image selection procedure described in Section
\ref{sec:simpleTextureMapping}.

Applying 2D alignments to this set of images works well for a number
of reasons. First, the nature of our input data and the selected
images is such that localization error chiefly occurs in two
dimensions, which correspond to the plane of the surface being
projected onto. This is because the backpack operator, during data
acquisition, makes efforts to walk within a few meters of, and
parallel to all walls being scanned. As a result, the translational
error of camera poses is quite minimal in the dimension perpendicular
to the wall surface being textured. Furthermore, because our set of
images has minimal distance from its corresponding wall surface,
rotational errors are less pronounced, except for rotation around the
axis perpendicular to the surface being textured. Therefore, most
translation and rotational errors occur in the 2 dimensions of the
wall surface's plane. This of course does not apply for ceilings and
floors; however, the lack of features on most ceilings and floors
means that 3D misalignment is not likely to visually make a
difference, beyond what can be done in 2D.

@@

After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we obtain a set
of image-based line segments by using Hough transforms to detect lines
in the image projections. These lines are detected in the image
projections rather than the original images, as the varying
orientation and distances of camera poses relative to surfaces results
in high variance of line lengths and strengths for real-world linear
features across the original images. We also gather a set of
geometry-based lines, which correspond to the lines comprising the
target surface's border, as well as lines formed where other surfaces
intersect the target surface. An example of these lines is shown in
red for a ceiling surface in Figure
\ref{fig:geometryAlignment}(a). Ideally, for perfect camera poses and
surface geometry, the lines in images corresponding to corners between
surfaces should match up exactly with corners in the 3D model. By
inducing such lines to match, we fit camera poses more accurately to
the surface, and to each other as well.

To align images to surface geometry, we collect pairs of image-based
and geometry-based line segments, which are within a distance and
angular difference threshold of each other. We have found a distance
threshold of 250 mm and an orientation difference threshold of
$10^\circ$ to work well for our datasets. For each pair of lines, we
compute the angular difference between the pair's image and geometry
lines. If there are 2 or more pairs with angular differences within
$1^\circ$, we select the two pairs with the longest noncollinear
image-based lines, and rotate the image such that the lines in the
pair with the longest image-based line become parallel. We then find a
translation such that the lines in that same pair overlap. This
translation has ambiguity in the dimension along the matched lines,
which is resolved by matching the midpoint of the image-based line in
the second pair to its corresponding geometry-based line. This is
shown in case (1) of Figure \ref{fig:geometryAlignment}(c). Thus, with
2 or more pairs, it is possible to obtain a fixed rotation and
translation, which are saved for usage in Section
\ref{sec:robustSIFTFeatureMatching}.

If there are not 2 or more pairs with similar angular differences, we
select the pair with the longest image-based line and apply a rotation
and the minimal translation to match its lines. This translation's
ambiguity however, can not be resolved, but is also saved to be used
in Section \ref{sec:robustSIFTFeatureMatching}. This is shown in case
(2) of Figure \ref{fig:geometryAlignment}(c). Finally, in the case
where there are no line pairs, we can still rotate images in order to
exploit patterns in indoor environments, as shown in case (3) of
Figure \ref{fig:geometryAlignment}(c). For instance, doors, windows,
furniture, etc. tend to have linear edges that are parallel to the
edges of the surfaces they are on. Similarly, lights, visible interior
walls, etc. which are visible in floor and ceiling images, tend to be
parallel to exterior walls corresponding to the edges of ceiling and
floor surfaces. Thus, we choose to minimize the angle between
image-based lines and geometry-based lines. We use the RANSAC
framework, with an inlier difference of $0.5^\circ$ in order to
successfully ignore outliers \cite{fischler1981random}.

@@

In order to correctly texture surfaces, it is important to detect and
remove parts of image projections containing texture for occluding
surfaces. For instance, in Figure \ref{fig:occlusion}(a), an image
used to texture the orange target surface also contains part of a gray
occluding surface. We remove this incorrect texture by recursively
performing ray-polygon intersection tests between the camera location
and every surface in our model except the target surface
\cite{rayintersection}. These intersection tests are performed at the
corners of a grid overlaid upon the target surface. Where all four
corners of a grid section are occluded, texture is removed. Where one
or more corners are occluded, the grid is subdivided into four, and
the process repeats. Occlusion checking works entirely with geometry,
so by ensuring that images match geometry using
\ref{sec:geometryAlignment}'s alignment procedure, texture belonging
to other surfaces is accurately removed, as seen in Figure
\ref{fig:occlusion}(b).

>>>>

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover its pose within an environment while simultaneously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment. Mounting such devices on a platform carried by
an ambulatory human provides unique advantages over static or
vehicular-based systems on wheels in terms of agility and portability,
but can also result in larger localization error
\cite{liu2010indoor}. As a result, common methods for texture mapping
generally produce poor results. In this thesis, we present an approach
to texture mapping 3D models of indoor environments in the presence of
uncertainty and noise in camera poses. In particular, we consider data
obtained from a human-operated backpack system, detailed in Section
\ref{sec:dataAcquisition}.

\begin{figure}
  \centering
  \includegraphics[width=6in]{flowchart.jpg}
  \caption{The proposed texture mapping procedure\\}
  \label{fig:flowchart}
\end{figure}

The overall block diagram for the proposed texture mapping procedure
is shown in Figure \ref{fig:flowchart}, where the number attached to
each box indicates the section in which the concept in the box is
explained in this thesis. First, the geometry of an environment is
split into regions, each of which will be textured independently and
in parallel. For each region, we begin by selecting a set of images
that spans the entire region's surface with high resolution
imagery. We then use recovered noisy camera poses to project these
selected images onto the surface. These projections are rotated and
translated in 2D, in order to maximally align them with the surface's
geometry, as well as to each other, allowing us to handle both errors
in geometry as well as camera poses. For surfaces and camera poses at
arbitrary locations or orientations, generally horizontal surfaces, we
propose a tile-based approach for sampling high-resolution portions of
images and compositing them into a texture. In cases where cameras
have consistently perpendicular viewing angles to the surfaces under
consideration, generally vertical surfaces, we demonstrate a superior
approach based on a shortest-path computation that leads to more
seamless textures.

The remainder of the thesis is organized as follows. Section
\ref{sec:dataAcquisition} provides background on the data acquisition
system and describes a region segmentation procedure. Section
\ref{sec:relatedWork} covers existing approaches to image stitching,
and their performance on our datasets. Section
\ref{sec:imageSelection} explains how to downsample the set of
available images by selecting those with the best orientation and
distance from surfaces under consideration. Section
\ref{sec:2dAlignment} contains the proposed approach towards 2D image
alignment, followed by Section \ref{sec:imageCompositing}, which
describes two methods of selecting and compositing images to create
the final texture. Sections \ref{sec:results} and \ref{sec:conclusion}
contain results and conclusions.

@@

This section contains background information describing the
backpack-mounted system and its operation. The hardware and operation
of the backpack system, as well as the postprocessing of acquired
data, play an important role in motivating the approaches described in
this thesis. The following two sections will focus on image and raw
data acquisition, and a method of partitioning recovered environment
geometry, respectively.

\subsection{Acquisition Hardware}

A CAD model of the backpack-mounted system is shown in Figure
\ref{fig:backpack}. This backpack system has a number of laser range
scanners as well as 2 cameras facing left and right, each equipped
with fisheye lenses reaching an approximately 180$^{\circ}$ field of
view each and taking photos at a rate of 5 Hz \cite{liu2010indoor}. In
order to collect data, an operator wears the backpack system and
traverses the environment to be scanned, walking in straight lines at
a standard pace and making an effort to walk parallel to all walls in
the area. Multiple localization and loop-closure algorithms are then
applied to the raw data collected by the onboard sensors
\cite{chen2010indoor, kua2012loopclosure, liu2010indoor}, and the
backpack is localized\footnote{In this thesis, we use the terms
  localization and pose recovery interchangeably, in that they both
  refer to recovering position and orientation.}  over its data
collection period. This involves recovering the 6 degrees of freedom
for the backpack itself as well as the cameras rigidly mounted on it.
These results are noisy, which motivates the approaches in this
thesis. Once this is complete, the data from the laser range scanners
is used to generate a 3D point cloud of the surrounding
environment. This point cloud is used to construct either a
low-resolution planar model, such as for usage in basic navigation or
energy analysis \cite{sanchez2012point, turnerfloorplan}, or a
high-resolution, detailed mesh, for applications such as accurate
remodeling or virtual walkthroughs \cite{turnerwatertight}.
Regardless of resolution, size, or detail, the model, consisting of
geometric surfaces in 3D space, along with the set of images captured
by the backpack's cameras and their noisy 3D poses, are the input to
the texture mapping problem.

\subsection{Geometry Partitioning}
\label{sec:geometryPartitioning}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=2.2in]{geomdeskwire_crop.png}}
  \subfloat[][]{\includegraphics[width=2.2in]{geomdeskregions_crop.png}}
  \subfloat[][]{\includegraphics[width=2.2in]{geomdesktexturetris_crop.png}}
  \caption{(a) Part of a model representing a desk and miscellaneous
    objects on top of the desk. (b) Triangles grouped into regions for
    texturing. (c) Final textured output.}
  \label{fig:deskGeom}
\end{figure}

In order to effectively texture a 3D model, we first divide the model
into a set of regions, each to be textured independently. The purpose
of this is to ensure texture continuity along large uniform areas,
while allowing texture boundaries to fall along natural environmental
boundaries. Because textures may not match at such boundaries where
two regions meet, it is important to minimize the visibility of
boundaries. This is done by encouraging region boundaries to occur at
large sharp corners, where texture continuity is not important. When
working with lower-resolution models, where surfaces generally
correspond to large planar features such as walls, ceilings, or
floors, and often meet at right angles, each planar surface in the
model can simply be treated as a distinct region and textured
separately.

With high-resolution models however, geometry often represents large
features as well as small details found in furniture, plants, and
other miscellaneous objects. For instance, in Figure
\ref{fig:deskGeom}(a), the geometry modeling each side of a desk as
well as objects on top of it consists of many small triangles at
various orientations. It is important that triangles belonging to the
same object or surface are textured together, as texturing them
independently might lead to visible misalignment between the textures
of each triangle, resulting in poor visualizations. On the other hand,
triangles belonging to different objects, such as the wall vs. objects
on the desk need not be textured together, as texture alignment
between the two is not important in terms of overall visual quality of
the scene. Thus, in this example, the triangles comprising each side
of the desk, objects on top of the desk, and the wall behind, should
all be grouped into separate regions. This region grouping is done by
first designating all contiguous coplanar groups of triangles as
different regions. Any regions that are less than $1 m^2$ in size are
then joined to their largest neighboring regions, as long as the angle
between them is under $90^{\circ}$. An example of such a grouping is
shown in Figure \ref{fig:deskGeom}(b). The same area with textures
successfully applied to each region is shown in Figure
\ref{fig:deskGeom}(c).

Though some of these regions from high-resolution models are not
completely flat, a plane is fitted to each region for the purposes of
texturing. This plane corresponds to the largest contiguous section of
coplanar triangles, as most regions are characterized by small
outcroppings from a largely planar surface, such as objects on a table
or features protruding from a wall. The 3D surface composing a region
is then projected onto its calculated plane, and the resulting 2D
planar polygon is used as the texturing surface. This 2D planar
approximation provides a simple, yet approximately accurate surface
onto which images can be projected in order to generate
textures. However, 3D surfaces are still retained in order to achieve
more accurate results when performing occlusion checks during the
texturing process. To improve runtime during such occlusion checks,
the triangles within all 3D surfaces are also loaded into a k-d tree
for quick lookup. Thus, for high-resolution models, each region has
both 3D geometry stored in a k-d tree for occlusion purposes, and an
approximate 2D representation stored as the texturing surface.

At this point, we either have a series of planar regions from a low
resolution model, or a set of planar approximations for regions as
well as an associated k-d tree from a high resolution model. The task
now is to generate a texture for each region.

\section{Related Work}
\label{sec:relatedWork}
There are many existing approaches to stitching together multiple
images to produce a larger, seamless image \cite{szeliski2006image,
  agarwalapanoramas, wangmultipleviews, coorg1997matching,
  debevechybrid, bernardinimultiplescans}. Generally, parts of images
are matched to each other, either through direct comparisons at a
pixel level, or through feature detection and matching. Images are
then transformed to maximize matches, often by computing homographies
between pairs of images, or by iteratively adjusting camera poses in 1
to 6 degrees of freedom.

@@

In prior work, we have experimented with integrating image stitching
with the iterative global localization algorithm used to recover the 6
degrees of freedom of the backpack acquisition system detailed in
Section \ref{sec:dataAcquisition} \cite{liu2010indoor}. When run on
long chains of images however, especially where features are sparse,
this approach produces distorted textures, as seen in Figure
\ref{fig:mosaic3D}(a). Furthermore, this approach is not closed-form,
and its iterative camera adjustment process over large datasets leads
to prohibitively long computation time.

The AutoStitch software package, based on research in the related area
of panorama genoration, also performs homography-based alignment, with
additional provisions that attempt to reduce drift and increase
overall efficiency \cite{panorama2d, autostitch}. Though AutoStitch
performs well in small areas, it often induces drift in images,
such that global alignment to the environment geometry is
lost. Furthermore, its performance relies on the presence of a dense
set of features, and it has trouble aligning wall sections with even
short segments of bare texture. The example in Figure
\ref{fig:mosaic3D}(b) was generated after many rounds of manual
tuning; we have empirically found that areas with fewer visual
features or repeating texture patterns simply failed outright with
AutoStitch.

@@

The geometry of the texture mapping process for a region, as described
in Section \ref{sec:geometryPartitioning}, is shown in Figure
\ref{fig:projection}. Given a set of images to texture a target
surface, camera matrix $P_i$ for the $i$th image transforms a 3D point
in the world coordinate system to a 2D point or pixel in image $i$'s
coordinates. A camera matrix $P_i$ is composed of the camera's
intrinsic parameters, containing focal length and image center, as
well as extrinsic parameters which specify the rotation and
translation of the camera's position in 3D world coordinates at the
time that image $i$ was taken. These extrinsic parameters are
determined by the backpack hardware and the corresponding localization
algorithms \cite{chen2010indoor, liu2010indoor, kua2012loopclosure}
and are noisy.

Since the backpack system takes pictures at a rate of 5 Hz, thousands
of images are available for texturing each surface in the 3D
model. Our objective in designing a texture mapping process is to
determine which of these images should be used, and where their
contents should map onto the final texture, in order to minimize any
visual discontinuities or seams that would suggest that the plane's
texture is not composed of a single continuous image. In the remainder
of this section, we propose an image subsampling procedure to obtain a
set of images for all further steps, and use it to derive a simple
tile-based texture mapping procedure.

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.8in]{Projection.pdf}
    \caption{Images are related to each surface through the camera
      matrices $P_{1..m}$. }
    \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.8in]{scoringFunction.jpg}
    \caption{Camera angle $\alpha$ and distance $d$ are minimized by
      maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
      \cdot \vec{n}$}
    \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

Our criteria for image subsampling has the following
stipulations. First, we want a set of images such that their
projections together cover up the entirety of our target surface, so
as to avoid holes in the final texture. Second, we desire our final
texture to have high resolution throughout; thus every location on the
target surface should have at least one image that contains high
resolution imagery for it. A straightforward way to accomplish these
goals is to discretize the target surface into small square tiles, and
for each tile, select the image that has the best viewing angle and
resolution for texturing that tile. In order to select the image that
can best texture a tile $t$, we first gather a list of candidate
images that contain all four of its corners; we can rapidly check this
by projecting $t$ into each image using the $P_i$ camera
matrices. Furthermore, each candidate image must have been taken at a
time when its camera had a clear line-of-sight to $t$, which can be
determined using standard ray-polygon intersection tests between the
camera location, $t$, and every other surface, or via the k-d tree
from Section \ref{sec:geometryPartitioning} if present
\cite{rayintersection}.

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since resolution decreases and camera pose errors become more
pronounced with distance, we wish to minimize the distance between
cameras and the surfaces they texture. Additionally, we desire images
that are projected perpendicularly, rather than obliquely, onto the
plane, maximizing the resolution and amount of useful texture
available in their projections, as well as minimizing any parallax
effects due to real-world geometry not accurately represented by the
digital 3D model. In other words, we wish to minimize the angle
between the tile's normal vector and the camera axis for images
selected for texturing that tile. These two criteria can be met by
maximizing the function $\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$
as shown in Figure \ref{fig:scoringFunction}, where $d$ is the
distance between the centers of a camera and a tile, and $\vec{n}$ and
$\vec{c}$ are the directions of the plane's normal and the camera axis
respectively.

When the images selected for each tile are used directly to texture
their respective tiles, image boundaries with abrupt discontinuities
between tiles are visible, as shown in Figure
\ref{fig:compareAll}(a). While it is clear that camera pose
inaccuracies are too severe for such a simple approach to work, the
selected images all appear to contain optimal camera angles with high
resolution, and much of their misalignment appears reconcilable using
2D transforms. This procedure generally selects around 10\% of the
possible images that could be used for texturing a surface, and not
only reduces the computational complexity of the remaining steps, but
also selects the most promising images for the remaining steps.

@@

\section{2D Image Alignment}
\label{sec:2dAlignment}
In this section, we describe our method for efficient and robust image
alignment. Rather than register all of our images in 3D, as many
state-of-the-art techniques for image stitching do, we instead align a
subset of images in 2D; this subset corresponds to the images selected
by the image selection procedure described in Section
\ref{sec:imageSelection}.

Applying 2D alignments to this set of images works well for a number
of reasons. First, the nature of our input data and the selected
images is such that localization error chiefly occurs in two
dimensions, which correspond to the plane of most surfaces being
projected onto. This is because the backpack operator, during data
acquisition, makes efforts to walk within a few meters of, and
parallel to all walls being scanned. Furthermore, the backpack system
maintains a fairly consistent distance from floors and ceilings during
data collection. As a result, the translational error of camera poses
is quite minimal in dimensions perpendicular to most surfaces being
textured, corresponding to, or parallel to walls, ceilings and
floors.

@@

After computing each image's projection onto the target surface, as
described in Section \ref{sec:imageSelection}, we can obtain a set of
image-based line segments by using Hough transforms to detect lines in
the image projections. These lines are detected in the image
projections rather than the original images, as the varying
orientation and distances of camera poses relative to surfaces results
in high variance of line lengths and strengths for real-world linear
features across the original images. We also gather a set of
geometry-based lines, which correspond to the lines comprising the
target surface's border, as well as lines formed where other surfaces
intersect the target surface. An example of these lines is shown in
red for a ceiling surface in Figure
\ref{fig:geometryAlignment}(a). Ideally, for perfect camera poses and
surface geometry, the lines in images corresponding to corners between
surfaces should match up exactly with corners in the 3D model. By
inducing such lines to match, we fit camera poses more accurately to
the surface, and to each other as well.

To align images to surface geometry, we collect pairs of image-based
and geometry-based line segments, which are within a distance and
angular difference threshold of each other. We have found a distance
threshold of 250 mm and an orientation difference threshold of
$10^\circ$ to work well for our datasets. For each pair of lines, we
compute the angular difference between the pair's image and geometry
lines. If there are 2 or more pairs with angular differences within
$1^\circ$, we select the two pairs with the longest noncollinear
image-based lines, and rotate the image such that the lines in the
pair with the longest image-based line become parallel. We then find a
translation such that the lines in that same pair overlap. This
translation has ambiguity in the dimension along the matched lines,
which is resolved by matching the midpoint of the image-based line in
the second pair to its corresponding geometry-based line. This is
shown in case (1) of Figure \ref{fig:geometryAlignment}(c). Thus, with
2 or more pairs, it is possible to obtain a fixed rotation and
translation for geometry alignment, which are saved for usage in
Section \ref{sec:robustSIFTFeatureMatching}.

If there are not 2 or more pairs with similar angular differences, we
select the pair with the longest image-based line, which corresponds
to a strong linear visual feature, and apply a rotation and the
minimal translation to match the pair's lines. This translation's
ambiguity however, can not be resolved, but is also saved to be used
in Section \ref{sec:robustSIFTFeatureMatching}. This is shown in case
(2) of Figure \ref{fig:geometryAlignment}(c). Finally, in the case
where there are no line pairs, we can still rotate images in order to
exploit patterns in indoor environments, as shown in case (3) of
Figure \ref{fig:geometryAlignment}(c). For instance, doors, windows,
furniture, etc. tend to have linear edges that are parallel to the
edges of the surfaces they are on. Similarly, lights, visible interior
walls, etc. which are visible in floor and ceiling images, tend to be
parallel to interior and exterior walls, corresponding to the
intersection lines and edges of ceiling and floor surfaces
respectively. Thus, we choose to minimize the angle between
image-based lines and geometry-based lines regardless of distance. We
use the RANSAC framework to compute a rotation angle that best
accomplishes this while ignoring outliers. \cite{fischler1981random}.

@@

In order to correctly texture surfaces, it is important to detect and
remove parts of image projections containing texture for occluding
surfaces. For instance, in Figure \ref{fig:occlusion}(a), an image
used to texture the orange target surface also contains part of a gray
occluding surface. We remove this incorrect texture by recursively
performing ray-polygon intersection tests between the camera location
and every surface in our model except the target surface
\cite{rayintersection}. If using planar approximations for regions, as
explained in Section \ref{sec:geometryPartitioning}, the k-d tree is
used instead. These intersection tests are performed at the corners of
a grid overlaid upon the target surface. Where all four corners of a
grid section are occluded, texture is removed. Where one or more
corners are occluded, the grid is subdivided into four, and the
process repeats. Occlusion checking works entirely with geometry, so
by ensuring that images match geometry using
\ref{sec:geometryAlignment}'s alignment procedure, texture belonging
to other surfaces is accurately removed, as seen in Figure
\ref{fig:occlusion}(b) vs. \ref{fig:occlusion}(c).
